---
title: "Stimuli Evaluation"
format:
  html:
    embed-resources: true
    echo: false
    warning: false
---

We ran a Prolific study and asked participants to rate the toxicity of 36 tweets (together with two attention checks). Out of a random sample of 100 tweets (see `validation-corpus.qmd`), each of the 40 participants drew 36 random tweets and rated their toxicity on a scale ranging from 0 (not toxic at all) to 100 (extremely toxic).

This document evaluates to which degree the human ratings are consistent (in itself) and with both the Perspective API score and the GPT-4o-mini score.


```{r packages_CRAN}
#| warning: false
# use rix at some point: https://github.com/ropensci/rix

options(repos = c(CRAN = "https://cloud.r-project.org"))

if (!requireNamespace("groundhog", quietly = TRUE)) {
  install.packages("groundhog")
}

pkgs <- c("data.table", "ggplot2", "patchwork", "stringr", "MOTE")

groundhog::groundhog.library(pkg = pkgs,
                             date = "2026-01-01") # R-4.4.1

rm(pkgs)
```

```{r set-seed}
set.seed(230691)
```

```{r constants}
HUMAN_THRESHOLD <- 0.6  # ratings above this (0-1 scale) count as toxic
```

```{r read_data}
raw <- fread(file = "../../data/raw/ratings_2026-04-22.csv")[valid_participant == 1]
time_spent <- fread(file = "../../data/raw/PageTimes-2026-04-22.csv")
```

```{r helpers}

a <- function(val, d = 3) MOTE::apa(val, decimals = d, leading = FALSE)

fmt_r <- function(x) {
  sprintf("r = %s, 95%% CI [%s, %s], t(%d) = %s, p %s",
          a(x$estimate),
          a(x$conf.int[1]),
          a(x$conf.int[2]),
          x$parameter,
          a(x$statistic, d = 2),
          ifelse(x$p.value < .001,
                 "< .001",
                 paste("=", a(x$p.value))))
}
```

```{r design}
c_primary <- "#a30d4e"

theme_paper <- theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor  = element_blank(),
    panel.grid.major  = element_line(colour = "grey92"),
    axis.title        = element_text(size = 10),
    plot.title        = element_text(face = "bold", size = 11),
    plot.subtitle     = element_text(colour = "#6a6a80", size = 9)
  )
```

## Exclusions and Data Preparation

```{r exclusions}
# Remove participants who failed attention checks
n_raw <- uniqueN(raw$participant_code)
raters <- raw[failed_attention == FALSE]
n_after <- uniqueN(raters$participant_code)

# Remove attention check posts and unrated rows
raters <- raters[!grepl("^attn_", doc_id)]
raters <- raters[!is.na(human_rating)]
```

## Aggregate Human Ratings

```{r aggregate}
tmp <- raters[,
              .(human_mean = mean(human_rating / 100),
                human_sd   = sd(human_rating / 100),
                n_raters   = .N),
              keyby = doc_id]

# Perspective and GPT scores are constant within doc_id; take first value
scores <- raters[,
                 .(perspective_toxicity = first(perspective_toxicity),
                   prop_human_toxic = mean(human_rating / 100 > HUMAN_THRESHOLD),
                   gpt_toxicity         = first(gpt_toxicity)),
                 keyby = doc_id]

agg <- merge(x = tmp, y = scores, by = "doc_id")

rm(tmp)
```

## Convergent Validity

```{r convergent-validity}
# Pearson r: human mean vs Perspective API
r_persp <- cor.test(agg$human_mean, agg$perspective_toxicity, method = "pearson")

# Pearson r: human mean vs GPT-4o-mini
r_gpt <- cor.test(agg$human_mean, agg$gpt_toxicity, method = "pearson")
```

Correlation between human averages and the Perspective API's score: `r fmt_r(r_persp)`.

Correlation between human averages and the GPT-4o-mini score: `r fmt_r(r_gpt)`.


```{r fig-convergent}
#| fig-width: 9
#| fig-height: 4.5

label_persp <- sprintf("r = %s [%s, %s]",
                       a(r_persp$estimate),
                       a(r_persp$conf.int[1]),
                       a(r_persp$conf.int[2]))

label_gpt <- sprintf("r = %s [%s, %s]",
                       a(r_gpt$estimate),
                       a(r_gpt$conf.int[1]),
                       a(r_gpt$conf.int[2]))

p1 <- ggplot(data = agg,
             mapping = aes(x = perspective_toxicity, y = human_mean)) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(method = "lm", se = TRUE, formula = "y ~ x",
              colour = c_primary, fill = c_primary, alpha = 0.12) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "#a0a0b8", linewidth = 0.4) +
  annotate("text", x = 0.05, y = 0.95, label = label_persp,
           hjust = 0, size = 3.2, colour = c_primary) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
  labs(title = "Human vs Perspective API",
       x = "Perspective toxicity score", y = "Mean human rating (0-1)") +
  theme_paper

p2 <- ggplot(data = agg,
             mapping = aes(x = gpt_toxicity, y = human_mean)) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(method = "lm", se = TRUE, formula = "y ~ x",
              colour = c_primary, fill = c_primary, alpha = 0.12) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "#a0a0b8", linewidth = 0.4) +
  annotate("text", x = 0.05, y = 0.95, label = label_gpt,
           hjust = 0, size = 3.2, colour = c_primary) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
  labs(title = "Human vs GPT-4o-mini",
       x = "GPT-4o-mini toxicity score", y = "Mean human rating (0-1)") +
  theme_paper

p1 + p2
```

## Perspective Score as Proportion Toxic

The Perspective API score is designed to approximate the share of people who would consider a post toxic. We operationalise the equivalent from our human ratings as the proportion of raters who assigned a score above 60 (mirroring our Perspectives API threshold). Both metrics are then on the same "share of people finding the post toxic" scale and can be compared directly.

```{r prop-toxic}
r_prop <- cor.test(agg$prop_human_toxic, agg$perspective_toxicity, method = "pearson")
```

Correlation between human proportions exceeding the threshold and the Perspective API's score: `r fmt_r(r_prop)`.


```{r fig-prop-toxic}
#| fig-width: 5.5
#| fig-height: 4.5

label_prop <- sprintf("r = %s [%s, %s]",
                      a(r_prop$estimate),
                      a(r_prop$conf.int[1]),
                      a(r_prop$conf.int[2]))

ggplot(data = agg,
       mapping = aes(x = perspective_toxicity, y = prop_human_toxic)) +
  geom_point(alpha = 0.55, size = 2) +
  geom_smooth(method = "lm", se = TRUE, formula = "y ~ x", 
              colour = c_primary, fill = c_primary, alpha = 0.12) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "#a0a0b8", linewidth = 0.4) +
  annotate("text", x = 0.05, y = 0.97, label = label_prop,
           hjust = 0, size = 3.2, colour = c_primary) +
  annotate("text", x = 0.72, y = 0.65, label = "identity line",
           hjust = 0, size = 2.8, colour = "#a0a0b8", angle = 38) +
  scale_x_continuous(breaks = seq(0, 1, .2),
                     labels = function(x) paste0(x * 10, "/10")) +
  scale_y_continuous(breaks = seq(0, 1, .2),
                     labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Perspective score vs share of raters finding post toxic",
       x = "Perspective score (out of 10 people)",
       y = "Share of raters rating > 0.6") +
  theme_paper
```

## Rating Distribution

```{r fig-distribution}
#| fig-width: 7
#| fig-height: 3.5

ggplot(agg, aes(x = human_mean)) +
  geom_histogram(binwidth = 0.05, fill = "#111118", colour = "white", linewidth = 0.3) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
  labs(title = "Distribution of mean human toxicity ratings",
       x = "Mean human rating (0-1)", y = "Number of posts") +
  theme_paper
```

## Ratings per Post

```{r fig-ratings-per-post}
#| fig-width: 9
#| fig-height: 3.5

agg_sorted <- agg[order(n_raters)]
agg_sorted[, rank := .I]

ggplot(agg_sorted, aes(x = rank, y = n_raters)) +
  geom_col(fill = "#111118", width = 0.8) +
  geom_hline(yintercept = 10, linetype = "dashed", colour = c_primary, linewidth = 0.5) +
  annotate("text", x = 1, y = 10.5, label = "minimum threshold (10)",
           hjust = 0, size = 3, colour = c_primary) +
  scale_y_continuous(breaks = seq(0, max(agg_sorted$n_raters) + 2, by = 2)) +
  labs(title = "Number of ratings per post (sorted ascending)",
       x = "Post rank (1 = fewest ratings)", y = "Number of ratings") +
  theme_paper

cat("Posts below 10 ratings:", agg_sorted[n_raters < 10, .N], "\n")
cat("Posts below  8 ratings:", agg_sorted[n_raters <  8, .N], "\n")
cat("Range:", min(agg_sorted$n_raters), "to", max(agg_sorted$n_raters), "\n")
```

# Attention Checks

```{r}
View(raw[failed_attention == TRUE, ])
```

