---
title: "Toxicity"

params:
  groundhogday: 2026-01-01

date: now
date-format: long
format: 
  html:
    embed-resources: true
    theme: cosmo
    code-line-numbers: true
toc: false
number-sections: false
fig-cap-location: top
execute:
  echo: true
  warning: false
---

# Steup

```{r install_packages}
#| warning: false
#| output: false

options(repos = c(CRAN = "https://cran.r-project.org")) 


if (!requireNamespace("groundhog", quietly = TRUE)) {
    install.packages("groundhog")
}

pkgs <- c("data.table", "stringr", "ggplot2", 
          "mgcv", "sandwich", "lmtest", "margins")

groundhog::groundhog.library(pkg = pkgs,
                             date = params$groundhogday)

rm(pkgs)
```

```{r}
set.seed(230690)
```

# Read, Manipulate, and Merge Data

```{r read_data}
# otree/DICE (remove the participant_code I used for testing)
dice <- fread(file = "../data/raw/DICE_2026-03-26.csv")[participant_code != "fdzno443" & session_code == "c1mufosc"]

# stimuli
stimuli <- fread(file = "../data/stimuli/toxic_movie_reactions.csv")

# qualtrics (maybe we need Jonas' API key, would be less ugly but also less reproducible though)
path <- "../data/raw/20260305_toxicity_pilot_v1_March 26, 2026_15.53.csv"
col_names <- names(fread(path, nrows = 0))
qualtrics <- fread(path, skip = 3, header = FALSE, col.names = col_names)
qualtrics <- qualtrics[grepl("^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}", StartDate)]
qualtrics[, (names(qualtrics)) := lapply(.SD, type.convert, as.is = TRUE)]
rm(path, col_names)
```

```{r rename_qualtrics}

setnames(x = qualtrics,
         old = c(
           "grewal-focal-brand_1","grewal-focal-brand_2","grewal-focal-brand_3",
           "grewal-focal-brand_4","grewal-focal-brand_5","grewal-focal-brand_6",
           "grewal-unrel-brand_1","grewal-unrel-brand_2","grewal-unrel-brand_3",
           "grewal-unrel-brand_4","grewal-unrel-brand_5","grewal-unrel-brand_6",
           "unaided-recall","aided-recall",
           "ad-involvement_1","ad-involvement_2","ad-involvement_3",
           "arousal-belanche_1","arousal-belanche_2","arousal-belanche_3",
           "info-diagnosticity_1","info-diagnosticity_2","info-diagnosticity_3",
           "brand-commitment_1","brand-commitment_2","brand-commitment_3",
           "choice-bernritter","purchase-bernritter_1",
           "Seen movie","movie-liking","brand-familiarity","social-media-usage"),
         new = c(
           "gfb_1","gfb_2","gfb_3","gfb_4","gfb_5","gfb_6",
           "gub_1","gub_2","gub_3","gub_4","gub_5","gub_6",
           "recall_text","recall_aided_raw",
           "ad_inv_1","ad_inv_2","ad_inv_3",
           "arousal_1","arousal_2","arousal_3",
           "info_diag_1","info_diag_2","info_diag_3",
           "comm_1","comm_2","comm_3",
           "choice","purchase",
           "seen_movie","movie_liking","brand_familiarity","social_media_usage"))
```

```{r composites}
# Brand attitudes
#     M&M's
qualtrics[, brand_attitude := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("gfb_", 1:6)]

#     Hershey's
qualtrics[, brand_attitude_c := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("gub_", 1:6)]

# Brand commitment (items 1 & 3 are reverse-coded on a 1–9 scale)
qualtrics[, comm_1r := 10 - comm_1]
qualtrics[, comm_3r := 10 - comm_3]
qualtrics[, brand_commitment := rowMeans(.SD, na.rm = TRUE),
          .SDcols = c("comm_1r","comm_2","comm_3r")]

# Mood
qualtrics[, mood_pos := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("mood_", 1:3)]
qualtrics[, mood_neg := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("mood_", 4:6)]

# Ad involvement, arousal, info diagnosticity (all 1–7, positive direction)
qualtrics[, ad_involvement := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("ad_inv_", 1:3)]
qualtrics[, ad_arousal := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("arousal_", 1:3)]
qualtrics[, info_diagnosticity := rowMeans(.SD, na.rm = TRUE),
          .SDcols = paste0("info_diag_", 1:3)]

# Attribution: brand / platform / algorithm responsibility (1–7 each)
qualtrics[, attribution := rowMeans(.SD, na.rm = TRUE), .SDcols = paste0("attribution_", 1:3)]

# Ad recall
#      unaided: regex "m&m" / "mms" etc.
qualtrics[, recall_unaided := as.integer(grepl(pattern = "m.?m", 
                                               x = recall_text, 
                                               ignore.case = TRUE))]

# aided: choice 1 in the multiple-choice list = M&M's
qualtrics[, recall_aided := as.integer(recall_aided_raw == 1)]

# Declaration: did participant notice the ad? (1 = Yes, 2 = No)
qualtrics[, noticed_ad := as.integer(declaration == 1)]

# Brand choice: 1 = M&M's
qualtrics[, choice_mms := as.integer(choice == 1)]
```

```{r merge_data}
merge_1 <- merge.data.table(x = dice, by.x = "participant_label",
                            y = qualtrics, by.y = "PID")

data    <- merge.data.table(x = merge_1,
                            y = stimuli,
                            by = "doc_id")

setorder(data, session_code, participant_code, displayed_sequence)
```

```{r toxicity_vars}
data[, experienced_toxicity := cumsum(toxicity), by = participant_code]

data[, binary_toxicity_05 := ifelse(test = toxicity >= 0.5, yes = 1, no = 0)]
data[, binary_toxicity_06 := ifelse(test = toxicity >= 0.6, yes = 1, no = 0)]
data[, binary_toxicity_07 := ifelse(test = toxicity >= 0.7, yes = 1, no = 0)]
data[, binary_toxicity_09 := ifelse(test = toxicity >= 0.9, yes = 1, no = 0)]

data[, feed_share_toxicity_05 := sum(binary_toxicity_05) / .N, by = participant_code]
data[, feed_share_toxicity_06 := sum(binary_toxicity_06) / .N, by = participant_code]
data[, feed_share_toxicity_07 := sum(binary_toxicity_07) / .N, by = participant_code]
data[, feed_share_toxicity_09 := sum(binary_toxicity_09) / .N, by = participant_code]
```

```{r delete_pii}
data[, participant_label := NULL]
data[, PROLIFIC_PID := NULL]
data[, IPAddress := NULL]
data[, condition.x := NULL]
data[, condition.y := NULL]
```


```{r}
ggplot(data = data[doc_id == 9999],
       mapping = aes(x = feed_share_toxicity_07)) +
  theme_minimal() +
  geom_histogram(bins = 30)
```


# Participant-Level Dataset

```{r part_level}
part_vars <- c(
  "participant_code", "session_code",
  "feed_toxicity", "time_on_feed","lottery_signup",
  "brand_attitude", "brand_attitude_c", "brand_commitment",
  "mood_pos", "mood_neg", "attribution",
  "ad_involvement", "ad_arousal", "info_diagnosticity",
  "recall_unaided", "recall_aided", "noticed_ad",
  "purchase", "choice_mms",
  "seen_movie", "movie_liking","brand_familiarity", "social_media_usage",
  "age", "gender", "ethnicity", "feed_share_toxicity_06"
)

part <- unique(data[, ..part_vars])

part[, experienced_toxicity := data[doc_id == 9999, 
                                    .(participant_code, experienced_toxicity)][part, on = "participant_code", experienced_toxicity]]

part[, scroll_depth := data[,
                            .(scroll_depth = {
                              reached <- displayed_sequence[!is.na(dwell_time)]
                              if (length(reached) > 0) max(reached) else 0}),
                            by = participant_code][part, on = "participant_code", scroll_depth]]

part[, ad_dwell_time := data[doc_id == 9999, 
                             .(participant_code, ad_dwell_time = focal_dwell_time)][part, on = "participant_code", ad_dwell_time]]
```

# Write Data

```{r write_data}
fwrite(x = data, file = "../data/processed/pilot_merged_data.csv")
fwrite(x = part, file = "../data/processed/pilot_participant_level_data.csv")
```

# Diagnostics

## Distribution of Toxicity

Is it problematic?

```{r}
ggplot(data = part,
       mapping = aes(x = feed_toxicity)) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)

ggplot(data = part,
       mapping = aes(x = experienced_toxicity/14)) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)
```


## Distribution of dwell times

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Vanilla dwell time on ad"
#|   - "Log dwell time on ad"

ggplot(data = part,
       mapping = aes(x = ad_dwell_time)) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)

ggplot(data = part,
       mapping = aes(x = log(ad_dwell_time))) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)
```

Does it make sense to divide these numbers by the amount of vertical pixels the ad was rendered in?

Anyways, here is the time_spent (on feed)

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Vanilla dwell time on feed"
#|   - "Log dwell time on feed"

ggplot(data = part,
       mapping = aes(x = time_on_feed)) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)

ggplot(data = part,
       mapping = aes(x = log(time_on_feed))) +
  theme_minimal() +
  geom_histogram(bins = 30, 
                 fill = "#1b7837", alpha = 0.5)
```


```{r}
#| eval: false
data[, unique(otf)]
```

Claude fasst das feedback wie folgt zusammen:

- Attentional capture: Most participants failed to notice the M&M ad, absorbed by Avatar post content.
- Habitual ad-avoidance: Several attributed non-noticing to automatic filtering developed through real-world social media use.
- Surprise at true focus: Participants assumed the study was about Avatar reactions; the ad reveal prompted amusement or mild frustration.
- Shallow processing among noticers: Brand recognition occurred, but impressions were weak; prompted recall was far more common than spontaneous recall.
- Pre-existing brand affect: A minority reported positive responses to the ad, tied to prior M&M preference rather than ad content.
- Methodological feedback:
  - Artificial, repetitive social media posts reduced ecological validity
  - Duplicate usernames posting conflicting opinions created confusion
  - Single-ad design may have increased target ad salience
  - Affective questions were difficult to answer for participants who had not consciously seen the ad
  
# Brand attitudes

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "M&M Brand Attitude"
#|   - "Hershey's Brand Attitude"

ggplot(data = part,
       mapping = aes(x = brand_attitude)) +
  theme_minimal() +
  geom_density(fill = "#1b7837", alpha = 0.5) +
  geom_vline(xintercept = part[, mean(brand_attitude)])

ggplot(data = part,
       mapping = aes(x = brand_attitude_c)) +
  theme_minimal() +
  geom_density(fill = "#762a83", alpha = 0.5) +
  geom_vline(xintercept = part[, mean(brand_attitude_c)])
```

```{r}
#| layout-ncol: 2
#| fig-cap: 
#|   - "M&M Brand Attitude"
#|   - "Hershey's Brand Attitude"

ggplot(data = part,
       mapping = aes(y = brand_attitude, x = experienced_toxicity)) +
  theme_minimal() +
  geom_point(col = "#1b7837", alpha = 0.5) +
  geom_smooth(formula = "y ~ x", method = "loess", se = TRUE)
  

ggplot(data = part,
       mapping = aes(y = brand_attitude_c, x = experienced_toxicity)) +
  theme_minimal() +
  geom_point(col = "#762a83", alpha = 0.5) +
  geom_smooth(formula = "y ~ x", method = "loess", se = TRUE)
  
```


## Recall & dwell time

```{r}
fit <- glm(formula = recall_aided ~ log(ad_dwell_time), 
            family="binomial",
            data = part)
```

```{r}
mar <- margins(model = fit, 
               variables = "ad_dwell_time")
summary(mar)

part_est <- as.data.frame(part[!is.na(ad_dwell_time) & !is.na(recall_aided)])

fit <- glm(recall_aided ~ log(ad_dwell_time), family = "binomial", data = part_est)

cplot(fit, x = "ad_dwell_time", what = "prediction",
      main = "Pr(Aided Recall) by Dwell Time")

cplot(fit, x = "ad_dwell_time", what = "effect",
      main = "AME of Dwell Time on Pr(Aided Recall)")
```

## Scroll depth

```{r}

```




# H1: Feed Engagement × Experienced Toxicity

Read two-lines test (Simonsohn 2018) which one can find [here](https://webstimate.org/twolines/twolines.R).
We need to decide on whether we use `feed_toxicity` or `experienced_toxicity` as a manipulation.
The latter feels more intuitive but also is somewhat endogenous. 

```{r source_twolines}
source("../analyses/twolines.R")
```

```{r h1_time}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Two Line Plot Uri"
#|   - "Two Line Plot ggplot"

tl_tof <- twolines(time_on_feed ~ feed_share_toxicity_06, data = as.data.frame(part), graph = 1)

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = time_on_feed)) +
  geom_point(alpha = 0.4) +
  geom_vline(xintercept = tl_tof$xc, linetype = "dashed", colour = "grey60") +
  geom_smooth(data = part[experienced_toxicity <= tl_tof$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#1b7837") +
  geom_smooth(data = part[experienced_toxicity >  tl_tof$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#762a83") +
  labs(x = "Experienced Toxicity", y = "Time on Feed (s)",
       title = "H1a: Time on Feed ~ Experienced Toxicity",
       subtitle = paste0("b1=", round(tl_tof$b1, 2), " p=", round(tl_tof$p1, 3),
                         "  |  b2=", round(tl_tof$b2, 2), " p=", round(tl_tof$p2, 3),
                         "  |  xc=", round(tl_tof$xc, 2))) +
  theme_minimal()
```

```{r h1_scroll}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Two Line Plot Uri"
#|   - "Two Line Plot ggplot"

tl_scroll <- twolines(scroll_depth ~ feed_share_toxicity_06, data = as.data.frame(part), graph = 1)

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = scroll_depth)) +
  geom_point(alpha = 0.4) +
  geom_vline(xintercept = tl_scroll$xc, linetype = "dashed", colour = "grey60") +
  geom_smooth(data = part[experienced_toxicity <= tl_scroll$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#1b7837") +
  geom_smooth(data = part[experienced_toxicity >  tl_scroll$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#762a83") +
  labs(x = "Experienced Toxicity", y = "Scroll Depth (prop. of feed)",
       title = "H1b: Scroll Depth ~ Experienced Toxicity",
       subtitle = paste0("b1=", round(tl_scroll$b1, 2), " p=", round(tl_scroll$p1, 3),
                         "  |  b2=", round(tl_scroll$b2, 2), " p=", round(tl_scroll$p2, 3),
                         "  |  xc=", round(tl_scroll$xc, 2))) +
  theme_minimal()
```

# H2: Ad Engagement × Experienced Toxicity

```{r h2_dwell}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Two Line Plot Uri"
#|   - "Two Line Plot ggplot"

# H2a: Ad dwell time (log-transformed for skew)
part[, log_ad_dwell := log(ad_dwell_time)]
tl_adt <- twolines(log_ad_dwell ~ experienced_toxicity, data = as.data.frame(part), graph = 1)

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = log_ad_dwell)) +
  geom_point(alpha = 0.4) +
  geom_vline(xintercept = tl_adt$xc, linetype = "dashed", colour = "grey60") +
  geom_smooth(data = part[experienced_toxicity <= tl_adt$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#1b7837") +
  geom_smooth(data = part[experienced_toxicity >  tl_adt$xc],
              formula = "y ~ x", method = "lm", se = TRUE, colour = "#762a83") +
  labs(x = "Experienced Toxicity", y = "log(1 + Ad Dwell Time)",
       title = "H2a: Ad Dwell Time ~ Experienced Toxicity",
       subtitle = paste0("b1=", round(tl_adt$b1, 2), " p=", round(tl_adt$p1, 3),
                         "  |  b2=", round(tl_adt$b2, 2), " p=", round(tl_adt$p2, 3),
                         "  |  xc=", round(tl_adt$xc, 2))) +
  theme_minimal()
```

```{r h2_ctr}
#| layout-ncol: 2
#| fig-cap: 
#|   - "Two Line Plot Uri"
#|   - "Two Line Plot ggplot"

# H2b: Ad CTR — logistic two-lines (binary outcome)
part[, ctr := as.numeric(lottery_signup)]
tl_ctr <- twolines(ctr ~ experienced_toxicity, data = as.data.frame(part),
                   link = "binomial", graph = 1)

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = ctr)) +
  geom_jitter(alpha = 0.3, height = 0.02) +
  geom_vline(xintercept = tl_ctr$xc, linetype = "dashed", colour = "grey60") +
  geom_smooth(data = part[experienced_toxicity <= tl_ctr$xc],
              method = "glm", method.args = list(family = "binomial"),
              se = TRUE, colour = "#1b7837") +
  geom_smooth(data = part[experienced_toxicity >  tl_ctr$xc],
              method = "glm", method.args = list(family = "binomial"),
              se = TRUE, colour = "#762a83") +
  labs(x = "Experienced Toxicity", y = "Ad Click (0/1)",
       title = "H2b: Ad CTR ~ Experienced Toxicity",
       subtitle = paste0("b1=", round(tl_ctr$b1, 2), " p=", round(tl_ctr$p1, 3),
                         "  |  b2=", round(tl_ctr$b2, 2), " p=", round(tl_ctr$p2, 3),
                         "  |  xc=", round(tl_ctr$xc, 2))) +
  theme_minimal()
```

```{r h2_recall}

# H2c: Unaided recall (logistic)
part[, recall_unaided_n := as.numeric(recall_unaided)]
tl_ur <- twolines(recall_unaided_n ~ experienced_toxicity, data = as.data.frame(part),
                  link = "binomial", graph = 0)

# H2d: Aided recall (logistic)
part[, recall_aided_n := as.numeric(recall_aided)]
tl_ar <- twolines(recall_aided_n ~ experienced_toxicity, data = as.data.frame(part),
                  link = "binomial", graph = 0)

recall_long <- melt(
  part[, .(participant_code, experienced_toxicity,
           Unaided = recall_unaided_n,
           Aided   = recall_aided_n)],
  id.vars       = c("participant_code","experienced_toxicity"),
  variable.name = "type",
  value.name    = "recalled"
)

ggplot(data = recall_long, 
       mapping = aes(x = experienced_toxicity, y = recalled, colour = type)) +
  geom_smooth(formula = "y ~ x", method = "loess", se = TRUE) +
  scale_colour_manual(values = c("#1b7837","#762a83")) +
  labs(x = "Experienced Toxicity", y = "P(Recall)", colour = NULL,
       title = "H2c–d: Ad Recall ~ Experienced Toxicity") +
  theme_minimal()
```

# RQ1: Brand Attitude × Experienced Toxicity

```{r rq1}
#| layout-ncol: 3

tl_ba <- twolines(brand_attitude ~ experienced_toxicity, data = as.data.frame(part), graph = 0)

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = brand_attitude)) +
  geom_point(alpha = 0.4) +
  geom_smooth(formula = "y ~ x", method = "loess", colour = "steelblue", se = TRUE) +
  labs(x = "Experienced Toxicity", y = "Brand Attitude — M&M's (1–7)",
       title = "RQ1: Brand Attitude ~ Experienced Toxicity") +
  theme_minimal()

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = purchase)) +
  geom_point(alpha = 0.4) +
  geom_smooth(formula = "y ~ x", method = "loess", colour = "steelblue", se = TRUE) +
  labs(x = "Experienced Toxicity", y = "Purchase Intent",
       title = "RQ1: Purchase Intent ~ Experienced Toxicity") +
  theme_minimal()

ggplot(data = part, 
       mapping = aes(x = experienced_toxicity, y = brand_commitment)) +
  geom_point(alpha = 0.4) +
  geom_smooth(formula = "y ~ x", method = "loess", colour = "steelblue", se = TRUE) +
  labs(x = "Experienced Toxicity", y = "Brand Commitment (1–9)",
       title = "RQ1: Brand Commitment ~ Experienced Toxicity") +
  theme_minimal()
```
